---
title: "Hands-on_Ex06"
author: "NeoYX"
editor: visual
execute: 
  freeze: auto
  warning: false
  #echo: false
  #message: false
  html:
    code-fold: True
    code-overflow: scroll
    code-summary: "Show the code"
    code-line-numbers: true
---

# 10  Visualising Uncertainty

## 10.1 Learning Outcome

## 10.2 Visualizing the uncertainty of point estimates

-   A point estimate is a single number, such as a mean.

Uncertainty is expressed as standard error, confidence interval, or credible interval

::: callout-important
Don't confuse the uncertainty of a point estimate (mean, median..) with the variation in the sample (standard deviation sigma, and variation sigma square etc)
:::

```{r}
pacman::p_load(tidyverse, plotly, crosstalk, DT, ggdist, gganimate)
```

-   [**DT**](https://rstudio.github.io/DT/) provides an R interface to the JavaScript library [DataTables](https://datatables.net/) that create interactive table on html page.

-   [Crosstalk](https://rstudio.github.io/crosstalk/) is an add-on to the htmlwidgets package. It extends htmlwidgets with a set of classes, functions, and conventions for implementing cross-widget interactions (currently, linked brushing and filtering).

-   [ggdist](https://mjskay.github.io/ggdist/) is an R package that provides a flexible set of `ggplot2` geoms and stats designed especially for visualizing distributions and uncertainty.

    Lets load the student exam data.csv

```{r}
exam <- read_csv('C:/yixin-neo/ISSS608-VAA/Hands-on_Ex/Hands-on_Ex05/data/Exam_data.csv')
```

### 10.2.1 Visualizing the uncertainty of point estimates: ggplot2 methods

The code chunk below performs the followings:

-   group the observation by RACE,

-   computes the count of observations, mean, standard deviation and standard error of Maths by RACE, and

-   save the output as a tibble data table called `my_sum`.

    ```{r}
    my_sum <- exam %>% 
      group_by(RACE) %>% 
      summarise(n=n(),
                mean=mean(MATHS),
                sd = sd(MATHS)) %>% 
      mutate(se=sd/sqrt(n-1))   #<<< standard error formula
    ```

Refer to lecture 4 slide 20 for mathematical formula explanation.

```{r}
knitr::kable(head(my_sum), format = 'html')
```

### 10.2.2 Visualizing the uncertainty of point estimates: ggplot2 methods

The code chunk below is used to reveal the **standard error** of **mean** maths score by race.

::: callout-note
Standard error is a measure of the variation of the mean of all the means from all samples of an underlying distribution.
:::

```{r}
ggplot(my_sum) +
  geom_errorbar(aes(x=RACE,
                    ymin=mean-se,
                    ymax=mean+se),
                width = 0.2,
                colour = 'black',
                alpha = 0.9,
                size=0.5) +
    geom_point(aes(x=RACE,
                 y=mean),
             stat = 'identity', #<<< actual points refer to mean 
             color='red', 
             size = 1.5,
             alpha = 1) + 
  ggtitle('Standard error of mean maths score by race')
```

::: callout-note
In the code above, **`stat = 'identity'`** means that the **`y`** values in the **`geom_point`** layer correspond to the **actual values** in the data frame, rather than a summary statistic like mean or median.
:::

### 10.2.3 Visualizing the uncertainty of point estimates: **ggplot2** methods

Lets plot a 95% confidence interval of mean maths score by race. The error bars should be sorted by the average maths scores. (Refer to take-home ex 1 on sorting by mean)

```{r}
ggplot(my_sum) +
  geom_errorbar(aes(x=reorder(RACE,-mean),
                    ymin=mean-se,
                    ymax=mean+se),
                width = 0.2,
                colour = 'black',
                alpha = 0.9,
                size=0.5) +
    geom_point(aes(x=RACE,
                 y=mean),
             stat = 'identity', #<<< actual points refer to mean 
             color='red', 
             size = 1.5,
             alpha = 1) + 
  ggtitle('Standard error of mean maths score by race')
```
