---
title: "In-class_Ex04"
author: "NeoYX"
date: '5 May 2023'
date-modified: "`r Sys.Date()`"
editor: visual
execute: 
  freeze: auto
  warning: false
  #echo: false
  #message: false
  html:
    code-fold: True
    code-overflow: scroll
    code-summary: "Show the code"
    code-line-numbers: true
---

We will load the following libraries:

-   rstatic : stats packages ready to go

-   tidyverse: will automatically provide read_r() \<- for read_csv()

```{r}
pacman::p_load(rstatix, gt, patchwork,tidyverse,nortest)
```

```{r}
exam <- read_csv('C:/yixin-neo/ISSS608-VAA/Hands-on_Ex/Hands-on_Ex05/data/Exam_data.csv')
```

```{r}
library(car)
b <- qqPlot(exam$ENGLISH)
b
```

```{r}
a <- ad.test(exam$ENGLISH)
```

HTML object

```{r}
shapiro.test(exam$ENGLISH)
```

tibble object

```{r}
exam %>% 
  shapiro_test(ENGLISH)
```

FROM Prof kam

recall hands-on 3 we use DT to create an interactive table , however it is not recognized by patchwork

```{r}
#| fig-width: 4
#| fig-height: 4

qq <- ggplot(exam,
       aes(sample=ENGLISH)) +  #<<< use a new argument call sample: el scores
  stat_qq() +
  stat_qq_line()

sw_t <- exam %>% 
  shapiro_test(ENGLISH) %>% gt()   #<<< make into a gt format (will give a nice table) if shapiro.test gives output in another output

tmp <- tempfile(fileext = '.png') # create  temp table
gtsave(sw_t, tmp)  # use gtsave() to save sw_t into tmp folder
table_png <- png::readPNG(tmp, native = TRUE)

qq+table_png

```
